Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-27 parallel-tractability-J-v9/response.tex

\begin{quote}
It would be much better if the ontology names corresponded to the example
numbers, i.e., the ontology in Example 6 was Oex6.
\end{quote}

We made the ontology subscripts correspond to the example numbers
for Example~5, Example~6, Example~7 and Example~9.

\begin{quote}
The notion of a simple concept is very much tied to the algorithm developed
in the paper.  How often do actual ontologies have constructs that could
make them intractable that are rended tractable by having simple concepts?
If there are none, then this class is uninteresting, particularly as the
paper makes such strong claims about worrying about particular ontologies.
This is different from an ontology only containing simple concepts.
\end{quote}

We investigated real ontologies from different data source. These ontologies cover
several domains and only contain simple concepts (these ontologies can be
obtained from the address given in the last paragraph of
Section~6.1).\\

@@

We used the speedup numbers to show different performances between the
ontologies in $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and
that not in $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.\\

In the revision, we provide the formula for the speedup and analyze
the speedup numbers.

\begin{quote}
There is a claim that the benchmarking distinguishes between the Ddhlo and
D-dhlo.  However, Figure 5 doesn't really show this well, if at all.   For
example, why should processing slow down with more threads at all?  Further,
it appears that even for several of the more well-behaved ontologies that there
is very little speedup between 6 and 8 processors.
\end{quote}

ParallelDHL is mainly
optimized based on the graph depth. Thus, when the graph depth is the
dominate factor that determines the reasoning time, the experiment 
should have a better performance. 
In the original experiments, the eight ontologies were used to make
the experiments close to real-world scenarios.
Since the graph depths of different ontologies affect the reasoning times
to different degrees, the experimental results are not very clear
to explain the issue of parallel tractability.
This is also why
Figure 5 does not show the differences between the ontologies in
$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ and those not in
$\mathcal{D}_{\textit{\text{dhl}}(\circ)}$ well.\\

In order to make the graph depth as a dominate factor, we modified the LUBM benchmark to 
generate ontologies of different graph depths and conducted the new experiments
in the revision.

### 2018-02-27 parallel-tractability-J-v9/appendix.tex
final polishing

\noindent\emph{Proof:} We first prove that the number of iterations of \ref{alg1:updateG}
is actually the depth of the constructed materialization graph.
We define the depths of nodes iteratively as follows:
for each of the original node $v$, \texttt{depth}($v$)=0;
for each implicit node $v'$ whose parents are $v_1,...,v_i$, \texttt{depth}($v'$)=$\max$$\{$ \texttt{depth}($v_1$), ..., \texttt{depth}($v_i$) $\}+1$.
By performing an induction on the number of iterations of \ref{alg1:updateG},
one can check that an implicit node $v'$ has to be added to $\mathcal{G}$ in the $n^{th}$ iteration
where $n$=\texttt{depth}($v'$).
Further, \texttt{depth}($\mathcal{G}$)=$\max$$\{$\texttt{depth}($v_i$)$|v_i$ is in $\mathcal{G}\}$.
We then have that the number of iterations of \ref{alg1:updateG} is \texttt{depth}($\mathcal{G}$).

@@

\emph{Inductive case}.
According to the induction hypothesis, the partial graph constructed after
the $i^{th}$ ($i>1$) iteration is a materialization graph, denoted by $\mathcal{G}^{i}$.
We have to show that the partial graph constructed after the $(i+1)^{th}$ iteration is
also a materialization graph, denoted by $\mathcal{G}^{i+1}$.
The partial graph $\mathcal{G}^{i+1}$ is updated in $\mathsf{A}_{\text{opt}}$ by performing Step $(\textbf{\romannumeral3})$
of Algorithm~$\mathsf{OPT}$.
Suppose that \texttt{rch}$(B_k,H)\in S_{\textit{\!\tiny rch}}$ is checked.
It corresponds to the rule instantiation $B_1,..,B_k,..,B_n\rightarrow H$.
Algorithm~$\mathsf{OPT}$ next checks that there exists a node $B'$ such that
\texttt{rch}$(B',B_k)\in S^*_{\!\tiny rch}$
and $B'$ is in $\mathcal{G}$. This means that $B_k$ and $H$ are derivable.
The algorithm then adds new nodes $B_k$ and $H$ to $\mathcal{G}^{i}$
according to three cases.
(1) If $H$ is not in $\mathcal{G}^{i}$ while $B_k$ is in $\mathcal{G}^{i}$,
then $H$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are created.
(2) If neither of $H$ and $B_k$ is in $\mathcal{G}^{i}$,
then $H$ and $B_k$ are added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are also created.
In the above two cases, $H$ is a new node. Thus,
$\mathcal{G}^{i+1}$ is acyclic.
(3) If $H$ is in $\mathcal{G}^{i}$ and $H$ has no parent,
$B_k$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),...,e(B_n, H)$ are also created.
For the case that $H$ is in $\mathcal{G}^{i}$ and $H$ has parents,
the algorithm does nothing.
Thus, $\mathcal{G}^{i+1}$ is acyclic and satisfies the definition of materialization graph.

@@

\emph{Inductive cases}. For each datalog rule of the form `$B_1,...,B_n\rightarrow H$'
that is rewritten from some DL-Lite$_\mathcal{R}$ axiom, we have the induction hypothesis
that $B_1,...,B_n$ have SWD paths.
We are left to prove that $H$ has an SWD path as well
in all materialization graphs of $\mathcal{O}$.

@@

\begin{enumerate}[leftmargin=12ex,label=Case~1.\arabic*]
\item $B\sqsubseteq A$. According to the induction hypothesis,
node $B(x)$ has an SWD path, denoted by $(v_1,...,v_n,B(x))$.
We have that, in some materialization graph, node $A(x)$
has an SWD path of the form $(v_1,...,v_n,B(x),A(x))$.

\item $\exists R\sqsubseteq A$. Node $R(x,y)$ has an SWD path,
denoted by $(v_1,...,v_n,R(x,y))$,
according to the induction hypothesis.
It is obvious that node $A(x)$
has an SWD path of the form $(v_1,...,v_n,B(x),R(x,y))$
in some materialization graph.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex,label=Case~2.\arabic*]
\item $S\sqsubseteq R$. According to the induction hypothesis,
node $S(x,y)$ has an SWD path, denoted by $(v_1,...,v_n,S(x,y))$.
Obviously, node $R(x,y)$
has an SWD path of the form $(v_1,...,v_n,S(x,y),R(x,y))$
in some materialization graph.

@@

    \item $R'$ is a non-simple role. By the induction hypothesis, if $R'(a,b)\in\delta_{R'}$ holds,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(a,b)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(a,c_1),R''(c_1,c_2),...,R''(c_n,b)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R$ and Definition~\ref{def:deltaRdhl},
        we have that $R(a,c_1),R(c_1,c_2),...,$ $R(c_n,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple role. By the induction hypothesis, if $R'(b,a)\in\delta_{R'}$,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(b,a)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(b,c_1),R''(c_1,c_2),...,R''(c_n,a)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R^-$ and Definition~\ref{def:deltaRdhl},
        we have that $R(a,c_n),R(c_n,c_{n-1}),...,$ $R(c_1,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

@@

In the above definition, for a role sequence $L=R_1R_2,...,R_n$, let $L^-=R_n^-,...,R_2^-R_1^-$.
We then give the following definition for $\delta_R$.

\begin{definition}\label{def:deltaRdhlplus}
Let $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$ be
a DHL$(\circ)$ ontology with $\textbf{RN}$ the set of role names. For each
$R\in\textbf{NR}$, we define $\delta_R$ as the following set of assertions:
\begin{align*}
\delta_R =\ & \{ R(a,b) \mid R(a,b) \in\mathcal{A} \}\ \cup\\
& \{ R(a,b) \mid R'(a,b)\in\mathcal{A} \text{ and } R' \sqsubseteq_{*}
  R \}\ \cup\\
& \{ R(a,b) \mid R'(b,a)\in\mathcal{A} \text{ and } R'\sqsubseteq_{*}
  R^- \}\\
& \{ R(a,b) \mid R_0R_1,\ldots,R_n\in\mathcal{L}(R),
  R_i(x_i,x_{i+1})\in\mathcal{A}, i\in\{0,1,...,n\}, x_0=a, x_{n+1}=b \}.
\end{align*}
% For each $R\in\textbf{RN}$, let $\delta_R$ be the set of all assertions as follows:
% \begin{enumerate}[leftmargin=4ex,label=\arabic*.]
% \item for each $R(a,b)\in\mathcal{A}$;
% \item $R(a,b)$, for each $R'(a,b)\in\mathcal{A}$ and $R'\sqsubseteq_{*}R$;
% \item $R(a,b)$, for each $R'(b,a)\in\mathcal{A}$ and $R'\sqsubseteq_{*}R^-$;
% \item $R(a,b)$, for each $R_0R_1,...,R_n\in\mathcal{L}(R)$, where $R_i(x_i,x_{i+1})\in\mathcal{A}$ and $x_0=a, x_{n+1}=b$ for $i\in\{0,1,...,n\}$.
% \end{enumerate}
\end{definition}

\begin{lemma}\label{lemma:tiplus}
Let
    $P=\langle R, \textbf{I}\rangle$ be the datalog program for a DHL$(\circ)$
    ontology
    $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$. We
  then have that
$P\models R(a,b)$ implies: (1) if $R$ is a simple role, $R(a,b)\in\delta_R$;
(2) if $R$ is a transitive role, then $R(a,b)\in\delta^*_R$;
(3) if $R$ is a non-simple but not transitive role, then there exists a role sequence $R_0R_1,...,R_n\in\mathcal{L}(R)$ such that
$R_i(x_i,x_{i+1})\in\delta_{R_i}$, or, $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$ if $R_i$ is a transitive role,
where $i\in\{0,1,...,n\}$ and $x_0=a, x_{n+1}=b$.
\end{lemma}

@@

\begin{enumerate}[leftmargin=12ex,label=Case~1.\arabic*, ref=Case~1.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$. Since $R$ is a simple role, $R'$ has to
    be a simple role according to Definition~\ref{def:deltaRdhlplus}. By the induction hypothesis,
    $R'(a,b)\in\delta_{R'}$ holds (Case~1.1.1). If $R'(a,b)\in\mathcal{A}$, $R(a,b)\in\delta_{R}$
    holds according to Definition~\ref{def:deltaRdhlplus} (Case~1.1.2). If there exists some
    role $R''$ such that $R''(a,b)\in\mathcal{A}$ and $R''\sqsubseteq R'$, we also have that
    $R(a,b)\in\delta_{R}$ because $R''\sqsubseteq_* R$ holds (Case~1.1.3). If there exists some
    role $R''$ such that $R''(b,a)\in\mathcal{A}$ and $R''\sqsubseteq R'^-$, the case is
    similar to Case~1.1.2 (Case~1.1.4). There exists $R_0R_1,...,R_n\in\mathcal{L}(R')$,
    where $R_i(x_i,x_{i+1})\in\mathcal{A}$ and $x_0=a, x_{n+1}=b$ for $i\in\{0,1,...,n\}$.
    According to Definition~\ref{def:language}, $R_0R_1,...,R_n\in\mathcal{L}(R)$ holds since
    $R'\sqsubseteq R\in\mathcal{R}$ (Case~1.1.5).\label{dhlplus:NTIcaseR1}

\item $R(a,b)$ is derived by applying the rule (R2), w.l.o.g., $R'(a,b)\rightarrow R(b,a)$,
    which also means that $R'\sqsubseteq R^-\in\mathcal{R}$. Since $R$ is a simple role, $R'$ has to
    be a simple role according to Definition~\ref{def:deltaRdhlplus}. By the induction hypothesis,
    $R'(b,a)\in\delta_{R'}$ (Case~1.2.1). If $R'(b,a)\in\mathcal{A}$, $R(a,b)\in\delta_{R}$
    holds according to Definition~\ref{def:deltaRdhlplus} (Case~1.2.2). If there exists some
    role $R''$ such that $R''(b,a)\in\mathcal{A}$ and $R''\sqsubseteq R'$, we also have that
    $R(a,b)\in\delta_{R}$ because $R''\sqsubseteq_* R^-$ holds (Case~1.2.3). If there exists some
    role $R''$ such that $R''(a,b)\in\mathcal{A}$ and $R''\sqsubseteq R'^-$, this case is
    similar to Case~1.2.2 (Case~1.2.4). There exists $R_0R_1,...,R_n\in\mathcal{L}(R')$,
    where $R_i(x_i,x_{i+1})\in\mathcal{A}$ and $x_0=b, x_{n+1}=a$ for $i\in\{0,1,...,n\}$.
    According to Definition~\ref{def:language}, $R_n^-,...,R_0^-\in\mathcal{L}(R)$ holds since
    $R'\sqsubseteq R^-\in\mathcal{R}$ (Case~1.2.5).\label{dhlplus:NTIcaseR2}

@@

    \item If there exists a role sequence
      $R_0R_1,...,R_m\in\mathcal{L}(R')$ s.t.\ $R_j(x_j,x_{j+1})\in\mathcal{A}$ and $x_0=c_i, x_{n+1}=c_{i+1}$ for $i\in\{0,1,...,m\}$
        and $R'\sqsubseteq_* R_i$, then $L_i=R_0R_1,...,R_m$ holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple but not a transitive role.
        By the induction hypothesis, there exists $R_0R_1,...,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,...,n\}$
        and $x_0=a, x_{n+1}=b$. Since $R'\sqsubseteq R$ holds, $R_0R_1,...,R_n\in\mathcal{L}(R)$ also holds.

    \item $R'$ is a transitive role.
        By the induction hypothesis, there exists $R'(a,b)\in\delta^*_{R'}$.
        Further, let $R'(a,c_1),R'(c_1,c_2),...,R'(c_n,b)\in\delta_{R'}$.
        Since $R'\sqsubseteq R\in\mathcal{R}$ holds, we also have that
        $R(a,c_1),R(c_1,c_2)$ $,...,R(c_n,b)\in\delta_{R}$. Obviously $R(a,b)\in\delta^*_{R}$ holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple but not transitive role. By the induction hypothesis, there exists $R_0R_1,...,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,...,n\}$
        and $x_0=b, x_{n+1}=a$. Since $R'\sqsubseteq R^-$ holds, $R_n^-,...,R_0^-\in\mathcal{L}(R)$ also holds.

     \item $R'$ is a transitive role.
        By the induction hypothesis, there exists $R'(b,a)\in\delta^*_{R'}$.
        Further, let $R'(b,c_1),R'(c_1,c_2),...,R'(c_n,a)\in\delta_{R'}$.
        Since $R'\sqsubseteq R^-\in\mathcal{R}$, we also have that
        $R(a,c_1),R(c_1,c_2)$ $,...,R(c_n,b)\in\delta_{R}$. Obviously $R(a,b)\in\delta^*_{R}$ holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple but not a transitive role.
        By the induction hypothesis, there exists $R_0R_1,...,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,...,n\}$
        and $x_0=a, x_{n+1}=b$. Since $R'\sqsubseteq R$, $R_0R_1,...,R_n\in\mathcal{L}(R)$ also holds.

@@

    \item $R'$ is a non-simple but not a transitive role. By the induction hypothesis, there exists $R_0R_1,...,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,...,n\}$
        and $x_0=b, x_{n+1}=a$. Since $R'\sqsubseteq R^-$, $R_n^-,...,R_0^-\in\mathcal{L}(R)$ also holds.

>>>>

\noindent\emph{Proof:} We first prove that the number of iterations of \ref{alg1:updateG}
is actually the depth of the constructed materialization graph.
We define the depths of nodes iteratively as follows:
for each of the original node $v$, \texttt{depth}($v$)=0;
for each implicit node $v'$ whose parents are $v_1,\ldots ,v_i$, \texttt{depth}($v'$)=$\max$$\{$ \texttt{depth}($v_1$), \ldots , \texttt{depth}($v_i$) $\}+1$.
By performing an induction on the number of iterations of \ref{alg1:updateG},
one can check that an implicit node $v'$ has to be added to $\mathcal{G}$ in the $n^{th}$ iteration
where $n$=\texttt{depth}($v'$).
Further, \texttt{depth}($\mathcal{G}$)=$\max$$\{$\texttt{depth}($v_i$)$|v_i$ is in $\mathcal{G}\}$.
We then have that the number of iterations of \ref{alg1:updateG} is \texttt{depth}($\mathcal{G}$).

@@

\emph{Inductive case}.
According to the induction hypothesis, the partial graph constructed after
the $i^{th}$ ($i>1$) iteration is a materialization graph, denoted by $\mathcal{G}^{i}$.
We have to show that the partial graph constructed after the $(i+1)^{th}$ iteration is
also a materialization graph, denoted by $\mathcal{G}^{i+1}$.
The partial graph $\mathcal{G}^{i+1}$ is updated in $\mathsf{A}_{\text{opt}}$ by performing Step $(\textbf{\romannumeral3})$
of Algorithm~$\mathsf{OPT}$.
Suppose that \texttt{rch}$(B_k,H)\in S_{\textit{\!\tiny rch}}$ is checked.
It corresponds to the rule instantiation $B_1,..,B_k,..,B_n\rightarrow H$.
Algorithm~$\mathsf{OPT}$ next checks that there exists a node $B'$ such that
\texttt{rch}$(B',B_k)\in S^*_{\!\tiny rch}$
and $B'$ is in $\mathcal{G}$. This means that $B_k$ and $H$ are derivable.
The algorithm then adds new nodes $B_k$ and $H$ to $\mathcal{G}^{i}$
according to three cases.
(1) If $H$ is not in $\mathcal{G}^{i}$ while $B_k$ is in $\mathcal{G}^{i}$,
then $H$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),\ldots ,e(B_n, H)$ are created.
(2) If neither of $H$ and $B_k$ is in $\mathcal{G}^{i}$,
then $H$ and $B_k$ are added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),\ldots ,e(B_n, H)$ are also created.
In the above two cases, $H$ is a new node. Thus,
$\mathcal{G}^{i+1}$ is acyclic.
(3) If $H$ is in $\mathcal{G}^{i}$ and $H$ has no parent,
$B_k$ is added to $\mathcal{G}^{i+1}$, and the edges $e(B_1, H),\ldots ,e(B_n, H)$ are also created.
For the case that $H$ is in $\mathcal{G}^{i}$ and $H$ has parents,
the algorithm does nothing.
Thus, $\mathcal{G}^{i+1}$ is acyclic and satisfies the definition of materialization graph.

@@

\emph{Inductive cases}. For each datalog rule of the form `$B_1,\ldots ,B_n\rightarrow H$'
that is rewritten from some DL-Lite$_\mathcal{R}$ axiom, we have the induction hypothesis
that $B_1,\ldots ,B_n$ have SWD paths.
We are left to prove that $H$ has an SWD path as well
in all materialization graphs of $\mathcal{O}$.

@@

\begin{enumerate}[leftmargin=12ex,label=Case~1.\arabic*]
\item $B\sqsubseteq A$. According to the induction hypothesis,
node $B(x)$ has an SWD path, denoted by $(v_1,\ldots ,v_n,B(x))$.
We have that, in some materialization graph, node $A(x)$
has an SWD path of the form $(v_1,\ldots ,v_n,B(x),A(x))$.

\item $\exists R\sqsubseteq A$. Node $R(x,y)$ has an SWD path,
denoted by $(v_1,\ldots ,v_n,R(x,y))$,
according to the induction hypothesis.
It is obvious that node $A(x)$
has an SWD path of the form $(v_1,\ldots ,v_n,B(x),R(x,y))$
in some materialization graph.
\end{enumerate}

@@

\begin{enumerate}[leftmargin=12ex,label=Case~2.\arabic*]
\item $S\sqsubseteq R$. According to the induction hypothesis,
node $S(x,y)$ has an SWD path, denoted by $(v_1,\ldots ,v_n,S(x,y))$.
Obviously, node $R(x,y)$
has an SWD path of the form $(v_1,\ldots ,v_n,S(x,y),R(x,y))$
in some materialization graph.

@@

    \item $R'$ is a non-simple role. By the induction hypothesis, if $R'(a,b)\in\delta_{R'}$ holds,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(a,b)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(a,c_1),R''(c_1,c_2),\ldots ,R''(c_n,b)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R$ and Definition~\ref{def:deltaRdhl},
        we have that $R(a,c_1)$, $R(c_1,c_2)$, \ldots, $R(c_n,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple role. By the induction hypothesis, if $R'(b,a)\in\delta_{R'}$,
        then $R(a,b)\in\delta_R$ also holds. On the other hand, if there exists a transitive role
        $R''$ (if $R'$ is a transitive role, then $R''\equiv R'$) such that
        $R''\sqsubseteq_* R'$ and $R''(b,a)\in\delta_{R''}^*$. Since $\delta_{R''}^*$ is
        the transitive closure of $\delta_{R''}$, we then have that there must
        exist such atoms ($R''(b,c_1),R''(c_1,c_2),\ldots ,R''(c_n,a)$) in $\delta_{R''}$.
        Further, due to $R''\sqsubseteq_* R^-$ and Definition~\ref{def:deltaRdhl},
        we have that $R(a,c_n)$, $R(c_n,c_{n-1})$, \ldots, $R(c_1,b)\in\delta_{R}$. Thus, $R(a,b)\in\delta_{R}^*$
        also holds.
    \end{enumerate}

@@

In the above definition, for a role sequence $L=R_1R_2,\ldots ,R_n$, let $L^-=R_n^-,\ldots ,R_2^-R_1^-$.
We then give the following definition for $\delta_R$.

\begin{definition}\label{def:deltaRdhlplus}
Let $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$ be
a DHL$(\circ)$ ontology with $\textbf{RN}$ the set of role names. For each
$R\in\textbf{NR}$, we define $\delta_R$ as the following set of assertions:
\begin{align*}
\delta_R =\ & \{ R(a,b) \mid R(a,b) \in\mathcal{A} \}\ \cup\\
& \{ R(a,b) \mid R'(a,b)\in\mathcal{A} \text{ and } R' \sqsubseteq_{*}
  R \}\ \cup\\
& \{ R(a,b) \mid R'(b,a)\in\mathcal{A} \text{ and } R'\sqsubseteq_{*}
  R^- \}\\
& \{ R(a,b) \mid R_0R_1,\ldots,R_n\in\mathcal{L}(R),
  R_i(x_i,x_{i+1})\in\mathcal{A}, i\in\{0,1,\ldots ,n\}, x_0=a, x_{n+1}=b \}.
\end{align*}
% For each $R\in\textbf{RN}$, let $\delta_R$ be the set of all assertions as follows:
% \begin{enumerate}[leftmargin=4ex,label=\arabic*.]
% \item for each $R(a,b)\in\mathcal{A}$;
% \item $R(a,b)$, for each $R'(a,b)\in\mathcal{A}$ and $R'\sqsubseteq_{*}R$;
% \item $R(a,b)$, for each $R'(b,a)\in\mathcal{A}$ and $R'\sqsubseteq_{*}R^-$;
% \item $R(a,b)$, for each $R_0R_1,\ldots ,R_n\in\mathcal{L}(R)$, where $R_i(x_i,x_{i+1})\in\mathcal{A}$ and $x_0=a, x_{n+1}=b$ for $i\in\{0,1,\ldots ,n\}$.
% \end{enumerate}
\end{definition}

\begin{lemma}\label{lemma:tiplus}
Let
    $P=\langle R, \textbf{I}\rangle$ be the datalog program for a DHL$(\circ)$
    ontology
    $\mathcal{O}=\langle\mathcal{T},\mathcal{R},\mathcal{A}\rangle$. We
  then have that
$P\models R(a,b)$ implies: (1) if $R$ is a simple role, $R(a,b)\in\delta_R$;
(2) if $R$ is a transitive role, then $R(a,b)\in\delta^*_R$;
(3) if $R$ is a non-simple but not transitive role, then there exists a role sequence $R_0R_1,\ldots ,R_n\in\mathcal{L}(R)$ such that
$R_i(x_i,x_{i+1})\in\delta_{R_i}$, or, $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$ if $R_i$ is a transitive role,
where $i\in\{0,1,\ldots ,n\}$ and $x_0=a, x_{n+1}=b$.
\end{lemma}

@@

\begin{enumerate}[leftmargin=12ex,label=Case~1.\arabic*, ref=Case~1.\arabic*]
\item $R(a,b)$ is derived by applying the rule (R1), w.l.o.g., $R'(a,b)\rightarrow R(a,b)$,
    which also means that $R'\sqsubseteq R\in\mathcal{R}$. Since $R$ is a simple role, $R'$ has to
    be a simple role according to Definition~\ref{def:deltaRdhlplus}. By the induction hypothesis,
    $R'(a,b)\in\delta_{R'}$ holds (Case~1.1.1). If $R'(a,b)\in\mathcal{A}$, $R(a,b)\in\delta_{R}$
    holds according to Definition~\ref{def:deltaRdhlplus} (Case~1.1.2). If there exists some
    role $R''$ such that $R''(a,b)\in\mathcal{A}$ and $R''\sqsubseteq R'$, we also have that
    $R(a,b)\in\delta_{R}$ because $R''\sqsubseteq_* R$ holds (Case~1.1.3). If there exists some
    role $R''$ such that $R''(b,a)\in\mathcal{A}$ and $R''\sqsubseteq R'^-$, the case is
    similar to Case~1.1.2 (Case~1.1.4). There exists $R_0R_1,\ldots ,R_n\in\mathcal{L}(R')$,
    where $R_i(x_i,x_{i+1})\in\mathcal{A}$ and $x_0=a, x_{n+1}=b$ for $i\in\{0,1,\ldots ,n\}$.
    According to Definition~\ref{def:language}, $R_0R_1,\ldots ,R_n\in\mathcal{L}(R)$ holds since
    $R'\sqsubseteq R\in\mathcal{R}$ (Case~1.1.5).\label{dhlplus:NTIcaseR1}

\item $R(a,b)$ is derived by applying the rule (R2), w.l.o.g., $R'(a,b)\rightarrow R(b,a)$,
    which also means that $R'\sqsubseteq R^-\in\mathcal{R}$. Since $R$ is a simple role, $R'$ has to
    be a simple role according to Definition~\ref{def:deltaRdhlplus}. By the induction hypothesis,
    $R'(b,a)\in\delta_{R'}$ (Case~1.2.1). If $R'(b,a)\in\mathcal{A}$, $R(a,b)\in\delta_{R}$
    holds according to Definition~\ref{def:deltaRdhlplus} (Case~1.2.2). If there exists some
    role $R''$ such that $R''(b,a)\in\mathcal{A}$ and $R''\sqsubseteq R'$, we also have that
    $R(a,b)\in\delta_{R}$ because $R''\sqsubseteq_* R^-$ holds (Case~1.2.3). If there exists some
    role $R''$ such that $R''(a,b)\in\mathcal{A}$ and $R''\sqsubseteq R'^-$, this case is
    similar to Case~1.2.2 (Case~1.2.4). There exists $R_0R_1,\ldots ,R_n\in\mathcal{L}(R')$,
    where $R_i(x_i,x_{i+1})\in\mathcal{A}$ and $x_0=b, x_{n+1}=a$ for $i\in\{0,1,\ldots ,n\}$.
    According to Definition~\ref{def:language}, $R_n^-,\ldots ,R_0^-\in\mathcal{L}(R)$ holds since
    $R'\sqsubseteq R^-\in\mathcal{R}$ (Case~1.2.5).\label{dhlplus:NTIcaseR2}

@@

    \item If there exists a role sequence
      $R_0R_1,\ldots ,R_m\in\mathcal{L}(R')$ s.t.\ $R_j(x_j,x_{j+1})\in\mathcal{A}$ and $x_0=c_i, x_{n+1}=c_{i+1}$ for $i\in\{0,1,\ldots ,m\}$
        and $R'\sqsubseteq_* R_i$, then $L_i=R_0R_1,\ldots ,R_m$ holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple but not a transitive role.
        By the induction hypothesis, there exists $R_0R_1,\ldots ,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,\ldots ,n\}$
        and $x_0=a, x_{n+1}=b$. Since $R'\sqsubseteq R$ holds, $R_0R_1,\ldots ,R_n\in\mathcal{L}(R)$ also holds.

    \item $R'$ is a transitive role.
        By the induction hypothesis, there exists $R'(a,b)\in\delta^*_{R'}$.
        Further, let $R'(a,c_1),R'(c_1,c_2),\ldots ,R'(c_n,b)\in\delta_{R'}$.
        Since $R'\sqsubseteq R\in\mathcal{R}$ holds, we also have that
        $R(a,c_1),R(c_1,c_2)$ $,\ldots ,R(c_n,b)\in\delta_{R}$. Obviously $R(a,b)\in\delta^*_{R}$ holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple but not transitive role. By the induction hypothesis, there exists $R_0R_1,\ldots ,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,\ldots ,n\}$
        and $x_0=b, x_{n+1}=a$. Since $R'\sqsubseteq R^-$ holds, $R_n^-,\ldots ,R_0^-\in\mathcal{L}(R)$ also holds.

     \item $R'$ is a transitive role.
        By the induction hypothesis, there exists $R'(b,a)\in\delta^*_{R'}$.
        Further, let $R'(b,c_1),R'(c_1,c_2),\ldots ,R'(c_n,a)\in\delta_{R'}$.
        Since $R'\sqsubseteq R^-\in\mathcal{R}$, we also have that
        $R(a,c_1),R(c_1,c_2)$ $,\ldots ,R(c_n,b)\in\delta_{R}$. Obviously $R(a,b)\in\delta^*_{R}$ holds.
    \end{enumerate}

@@

    \item $R'$ is a non-simple but not a transitive role.
        By the induction hypothesis, there exists $R_0R_1,\ldots ,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,\ldots ,n\}$
        and $x_0=a, x_{n+1}=b$. Since $R'\sqsubseteq R$, $R_0R_1,\ldots ,R_n\in\mathcal{L}(R)$ also holds.

@@

    \item $R'$ is a non-simple but not a transitive role. By the induction hypothesis, there exists $R_0R_1,\ldots ,R_n\in\mathcal{L}(R')$
        such that $R_i(x_i,x_{i+1})\in\delta_{R_i}$ or $R_i(x_i,x_{i+1})\in\delta^*_{R_i}$, where $i\in\{0,1,\ldots ,n\}$
        and $x_0=b, x_{n+1}=a$. Since $R'\sqsubseteq R^-$, $R_n^-,\ldots ,R_0^-\in\mathcal{L}(R)$ also holds.
